---
title: "Parkinson’s Disease Analysis"
author: "Claire Green"
date: "28/02/2018"
output: html_document
---

#Data Sets

Datasets were searched for using the same filters as before - must be human tissue that expressed lewy bodies, must have at least 3 patients and 3 controls, and must...

10 datasets were identified:

| CodeID | Tissue/Cell type                                         | Variant       | Samples                                                                                                                                                 | Type of array                               | GEO Ref  | Comments                                                               | PIs                                                                                               |
|--------|----------------------------------------------------------|---------------|---------------------------------------------------------------------------------------------------------------------------------------------------------|---------------------------------------------|----------|------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------|
| DIJ    | Substantia Nigra                                         | not specified | 15 disease, 8 control                                                                                                                                   | Affymetrix Human Genome U133 Plus 2.0 Array | GSE49036 | Has some Braak staging                                                 | Dijkstra AA, Ingrassia AI, de Menezes RX, van Kesteren RE, Rozemuller AM, Heutink P, vandeBerg WJ |
| FFR    | Substantia Nigra                                         | not specified | 9 replicates for the controls and 16 replicates for the Parkinson's disease patients                                                                    | Affymetrix Human Genome U133 Plus 2.0 Array | GSE7621  |                                                                        | Ffrench-Mullen JM                                                                                 |
| MOR.SN | substantia nigra, split into medial and lateral portions | sporadic      | 15 samples of medial parkinsonian SN, 9 samples of lateral parkinsonian SN (24), 8 medial nigra control samples and 7 lateral nigra control samples(15) | Affymetrix Human Genome U133A Array         | GSE8397  | Age and sex                                                            | Moran LB, Graeber MB                                                                              |
| MID1   | Substantia Nigra pars compacta                           | not specified | 10 PD brain samples and 8 control                                                                                                                       | Affymetrix Human Genome U133 Plus 2.0 Array | GSE20141 |                                                                        | Middleton FA, Kim PD, Zhang-James Y, Davis RL                                                     |
| MID2   | whole substantia nigra                                   | not specified | 18 controls, 11 patients                                                                                                                                | Affymetrix Human Genome U133A Array         | GSE20292 | Age and sex                                                            | Middleton FA, James M, Zhang Y, Davis RL                                                          |
| DUM    | Cortex (BA9)                                             | not specified | 29 PD, 44 neurologically normal controls                                                                                                                | Illumina HiSeq 2000                         | GSE68719 | Have age of death, sex, and some samples have corresponding proteomics | Dumitriu A, Golji J, Labadorf AT, Gao B, Beach TG, Myers RH, Longo KA, Latourelle JC              |
| MOR.FC | Frontal cortex                                           | sporadic      | 3 Controls 5 Patients                                                                                                                                   | Affymetrix Human Genome U133A Array         | GSE8397  | Age and sex                                                            | Moran LB, Graeber MB                                                                              |
| LEW    | two regions of the medulla                               | not specified | 14 PD samples and 8 controls (two regions)                                                                                                                 | Affymetrix Human Genome U133A 2.0 Array     | GSE19587 | Age and sex                                                            | lewandowski nm, small sa                                                                          |
| MID3   | prefrontal cortex                                        | not specified | 15 controls 14 patients                                                                                                                                 | Affymetrix Human Genome U133A Array         | GSE20168 | Age and sex                                                            | Middleton FA, James M, Zhang Y, Davis RL                                                          |
| MID4   | putamen                                                  | not specified | 20 controls and  15 patients                                                                                                                            | Affymetrix Human Genome U133A Array         | GSE20291 | Age and sex                                                            | Middleton FA, James M, Zhang Y, Davis RL                                                          |

#Quality Control - PD_QC.R

Checked data distribution using PCA and boxplots. Some of the datasets have quite mixed samples wereas others separate quite nicely into controls and patients. Boxplots didn't suggest any samples were outliers.


```{r, out.width = "900px", out.height = "700px", echo=F}
knitr::include_graphics("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/QC/QC_plots.pdf")
```

#Differential Gene Expression

As with the TDP-43 data, microarray data was analysed using Limma. The genes were divided into positive and negative fold change. Intersected gene lists can be found in /Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/FoldChange. 

## Commonly Upregulated Genes

| ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~  | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~  | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~   | ~~~~~~~~~~~ | ~~~~~~~~~~~ |
|-------------|-------------|--------------|-------------|-------------|--------------|-------------|-------------|---------------|-------------|-------------|
| A4GNT       | C3          | CPED1        | FAM46A      | HCK         | KDM4A        | LY6G5C      | NRDE2       | RBMS2         | SMIM14      | TNFRSF10B   |
| ABCC6       | C3AR1       | CPM          | FAM60A      | HDC         | KDM4C        | LYVE1       | NYX         | RCC1          | SNAP23      | TNFRSF10D   |
| AC012065.7  | C8orf60     | CPSF3L       | FANCE       | HERC2P3     | KIAA0101     | LYZL6       | OR1F1       | RDH16         | SND1-IT1    | TNFRSF25    |
| ACP5        | C9orf3      | CROCC        | FAS         | HIST1H1T    | KIAA0226L    | MAF         | ORM1        | RHBDF2        | SOGA1       | TNFSF14     |
| ACSBG2      | CALML4      | CROCCP3      | FASLG       | HIST1H2AB   | KIAA0485     | MAFF        | OSM         | RIBC2         | SORBS1      | TNN         |
| ACSS3       | CASP1       | CSPG4        | FCGBP       | HIST1H2AD   | KIAA1614     | MAP2K7      | P4HA1       | RIPK2         | SOS2        | TP73-AS1    |
| ADH1B       | CASP2       | CTBP2        | FCGR2A      | HMOX1       | KIAA1661     | MAPKBP1     | PAGE1       | RNASE6        | SOX12       | TRAF1       |
| ADIPOQ      | CASP4       | CTD-2269F5.1 | FCGR2B      | HNRNPM      | KIF13A       | MAVS        | PAPOLA      | RP11-217B7.2  | SOX2        | TRAPPC10    |
| ADORA3      | CASP6       | CXCR4        | FCRL2       | HOXB8       | KLHL36       | MBD5        | PARP16      | RP11-255C15.3 | SOX9        | TRIM21      |
| AF007147    | CASP7       | CXorf21      | FGF17       | HP1BP3      | KLK2         | MBTD1       | PAWR        | RP11-403P17.4 | SP140L      | TRIM5       |
| AGAP10      | CASP8       | CXorf36      | FGF3        | HS3ST1      | KRI1         | MCM3AP-AS1  | PCDH12      | RP11-548H18.2 | SPG21       | TTLL5       |
| AGRP        | CATSPER2    | CYP21A1P     | FGFR1       | HSPA1A      | KRT19P2      | MECOM       | PCDHGA8     | RPA4          | SPINK1      | TUBD1       |
| ALOX5       | CCDC101     | CYP39A1      | FKBP10      | HSPA1L      | L2HGDH       | MED13L      | PDLIM1      | RPL23AP32     | SPTLC3      | TULP2       |
| ALOX5AP     | CCDC15      | CYTH4        | FLCN        | HSPA6       | LAG3         | METTL4      | PELI2       | RPL23AP53     | SRP19       | TYMP        |
| ALPK1       | CCDC40      | DAPP1        | FLJ21369    | HSPD1       | LAMB2        | METTL7A     | PGF         | RREB1         | SRSF5       | TYROBP      |
| AMELX       | CCL17       | DBF4B        | FLT1        | IFNA2       | LAT2         | MFAP5       | PGPEP1      | RRNAD1        | SSTR3       | UGGT1       |
| AMELY       | CCL22       | DCLRE1C      | FLT3LG      | IFNA21      | LECT1        | MICB        | PIDD1       | RUNX1-IT1     | ST20        | UPK3B       |
| AMH         | CCL27       | DDR1-AS1     | FMO5        | IGF1R       | LEF1         | MID1        | PIGO        | RUNX3         | ST6GAL1     | USP34       |
| ANGPT2      | CD14        | DENND2D      | FNDC3B      | IGFBP5      | LENEP        | MILR1       | PILRA       | S100A11P1     | STAG3L3     | VPS54       |
| ANP32A-IT1  | CD163       | DFFB         | FOXD1       | IGKV1-17    | LEPREL1      | MIR612      | PIM1        | S100A4        | STIP1       | VSIG4       |
| AP1G2       | CD207       | DIP2A        | FOXL2       | IGLL3P      | LEPREL4      | MKRN4P      | PLAC8       | SAFB2         | STK10       | WBSCR16     |
| APOBEC3C    | CD22        | DISC1        | FPR1        | IGSF9B      | LGALS9       | MS4A6A      | PLEKHF2     | SAT1          | STOM        | WDR4        |
| APOC2       | CD247       | DKFZP434A062 | FYB         | IL12RB1     | LILRA2       | MSH5        | PLGLB1      | SCAF4         | TAF4        | WDR52       |
| AQP8        | CD300A      | DMC1         | FZD5        | IL16        | LINC00115    | MSX1        | PLIN2       | SCD5          | TAL1        | WDR55       |
| ARHGAP25    | CD37        | DNA2         | FZD7        | IL17RB      | LINC00260    | MT1M        | PLK3        | SCIN          | TAS2R10     | WDR78       |
| ARHGDIB     | CD48        | DNAJB1       | FZD9        | IL18        | LINC00894    | MYH7B       | PLOD2       | SCNN1B        | TAS2R13     | XRCC2       |
| ARHGEF4     | CD68        | DNAJB6       | G0S2        | IL1R1       | LINC00963    | MYO1C       | PLTP        | SDCCAG3       | TAZ         | ZBED2       |
| ARID3A      | CD84        | DNASE2       | G3BP1       | IL21        | LLGL2        | MZT2B       | PODNL1      | SERPINA1      | TBX6        | ZC3HAV1     |
| ATAD2B      | CDA         | DOCK10       | GABPA       | INE1        | LOC100272216 | NABP1       | POGLUT1     | SERPINH1      | TBXAS1      | ZFP36L1     |
| ATHL1       | CDC14A      | DOCK2        | GAL3ST4     | INPP5B      | LOC100506699 | NACAP1      | POGZ        | SERTAD3       | TCF3        | ZMYND8      |
| ATP8B1      | CDK2AP2     | DSE          | GAS1        | INPP5D      | LOC101928198 | NBEAL2      | PPP1R14D    | SH2B2         | TFAP2B      | ZNF14       |
| ATXN7       | CEBPD       | DTYMK        | GATA1       | INVS        | LOC101928274 | NBR2        | PPP2R1B     | SHMT1         | TFEC        | ZNF217      |
| AVIL        | CEL         | EEF1DP5      | GCKR        | IRF4        | LOC101929240 | NCKAP1L     | PRR11       | SIGLEC7       | TFPI        | ZNF224      |
| AZGP1       | CHD2        | EFNA4        | GDF15       | IRF6        | LOC101929889 | NEUROG1     | PRRG4       | SIGLEC8       | TGFBR3      | ZNF235      |
| AZGP1P1     | CHKB-CPT1B  | EGF          | GDF9        | IRF7        | LOC102724229 | NFATC3      | PSCA        | SIM2          | TGIF1       | ZNF34       |
| AZU1        | CHRNG       | ELF4         | GEM         | ITFG2       | LOC102725016 | NFE2        | PSG11       | SIX6          | TIMM44      | ZNF354A     |
| BANP        | CHST4       | EMC10        | GJA9-MYCBP  | ITGB6       | LOC284009    | NHLH1       | PTPN14      | SLA           | TLR5        | ZNF430      |
| BC069804    | CLEC2B      | ENTPD1       | GJD2        | ITPR3       | LOC729164    | NOTCH2NL    | PTPRCAP     | SLAMF8        | TM4SF1      | ZNF446      |
| BMP8B       | CLIC2       | ENTPD7       | GK2         | JMJD6       | LOC79999     | NOX3        | RAB13       | SLC11A1       | TMEM176A    | ZNF516      |
| BRPF1       | CLK1        | EP300        | GP9         | KCNE4       | LPAR2        | NPIPA1      | RAC2        | SLC27A3       | TMEM19      | ZNF665      |
| BTBD18      | COL16A1     | EPOR         | GPR25       | KCNJ8       | LPAR4        | NPIPB15     | RAD52       | SLC2A5        | TMEM51      | ZNF670      |
| C10orf12    | COL5A1      | ETV1         | GRAMD3      | KCNK5       | LPP          | NPL         | RAD54L      | SLC35C2       | TMPRSS11E   | ZNF710      |
| C11orf16    | COL7A1      | FAM106A      | GTSE1       | KCNQ1       | LRRC1        | NPR3        | RBL1        | SLC7A9        | TMPRSS5     | ZNF721      |
| C1QB        | COL9A1      | FAM118A      | GUSBP3      | KCTD5       | LRRFIP1      | NR6A1       | RBM38       | SMAD6         | TNFAIP8     | ZNF74       |
|             |             |              |             |             |              |             |             |               |             | ZNF783      |


## Commonly Downregulated Genes
| ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~  | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ | ~~~~~~~~~~~ |
|-------------|-------------|-------------|-------------|--------------|-------------|-------------|-------------|-------------|-------------|-------------|
| ABCA11P     | ATP5D       | CLINT1      | EIF2B3      | GPN1         | LOC728392   | MZT2A       | PFDN2       | PSMD8       | SLC9A1      | TSSC1       |
| ABCF1       | ATP5F1      | CLIP3       | EIF2S1      | GPX4         | LRBA        | NAPA        | PFKM        | PSME3       | SLITRK5     | TTC19       |
| ABHD11      | ATP5G1      | CLTB        | EIF3F       | GRAMD1B      | LRPPRC      | NDEL1       | PFN2        | PSMG1       | SNCA        | TTC9        |
| ACAT1       | ATP6AP1     | CMAS        | EIF3I       | GSS          | LYRM4       | NDFIP1      | PGAM1       | PTBP2       | SNRPN       | TUBA1B      |
| ACAT2       | ATP6V0B     | COA3        | EIF3K       | GSTA4        | LYRM9       | NDUFA10     | PGRMC1      | PTDSS1      | SNX24       | TUBA1C      |
| ACP2        | ATP6V0C     | COA7        | EIF6        | GSTO1        | MAGED1      | NDUFA13     | PHB2        | PTP4A1      | SORD        | TUBB        |
| ACTB        | ATP6V1E1    | COMMD9      | ELOVL6      | HARS         | MAGED2      | NDUFA3      | PINK1       | PTPRA       | SPAG16      | TUBB2A      |
| ACTR1A      | ATP6V1F     | COPS3       | ELP3        | HERC6        | MAK16       | NDUFA7      | PITPNA      | RAB14       | SPAG7       | TUBB3       |
| ADAM23      | ATP6V1G2    | COPS4       | EMC7        | HINT1        | MAP1B       | NDUFA9      | PITPNB      | RABEPK      | SPCS1       | TUBG1       |
| AFG3L2      | ATP6V1H     | COPS5       | ENDOD1      | HLTF         | MAP1LC3B    | NDUFAB1     | PITRM1      | RAN         | SPCS3       | TUFM        |
| AGTPBP1     | AURKAIP1    | COPS7A      | ENOPH1      | HMCES        | MAPK10      | NDUFAF1     | PKI55       | RANGRF      | SPRYD7      | TUSC2       |
| AHCY        | AVPI1       | COQ3        | ENSA        | HMGN4        | MDH1        | NDUFB5      | PLD3        | RARS        | SRPRB       | TXNDC15     |
| AIFM1       | B3GNT2      | COX7A2L     | ENTPD6      | HNRNPA0      | MDH2        | NDUFS1      | PLEKHB2     | RBFOX2      | SRRD        | UBB         |
| AIMP2       | BABAM1      | COX8A       | ERCC1       | HPCAL4       | ME2         | NDUFS3      | PMPCA       | RBM3        | SSSCA1      | UBE2E3      |
| AJAP1       | BECN1       | CREB3       | ERP29       | HSPA12A      | ME3         | NDUFS5      | PMS2        | RITA1       | ST3GAL5     | UBE3B       |
| AK055981    | BEND5       | CRMP1       | EXOSC9      | IARS         | MECR        | NDUFV1      | PMS2P1      | RNASEH1     | STAT4       | UCHL1       |
| AK5         | BEX4        | CRY2        | FABP3       | IARS2        | MFN2        | NDUFV2      | PMS2P5      | RNF10       | STRAP       | UQCRC1      |
| AKAP12      | BFSP1       | CSNK2A1     | FAF1        | IDH3B        | MGST3       | NEDD8       | POLR2C      | RPA2        | STS         | UQCRFS1     |
| AKAP6       | BPGM        | CSRNP2      | FAM127A     | IDH3G        | MICU1       | NEU1        | POLR3C      | RPH3A       | STX12       | UQCRQ       |
| AKTIP       | BRE         | CTNNA2      | FAM134C     | IDI1         | MIEF1       | NGFRAP1     | POP4        | RPP14       | STXBP1      | UROD        |
| ALAS1       | BRF2        | CUL2        | FAM162A     | IMP4         | MIF         | NHP2        | PPFIA2      | RPS6KC1     | SUMO3       | UROS        |
| ALDH1A1     | BSN         | CX3CL1      | FAM168A     | INSIG1       | MIR21       | NIF3L1      | PPIA        | RRAGA       | SYN1        | USP7        |
| ANXA2       | BTBD3       | CXorf40A    | FAM206A     | IQSEC1       | MIR3656     | NIPSNAP1    | PPME1       | RTCB        | SYT11       | UTP18       |
| ANXA6       | C11orf49    | CXorf40A    | FBXW2       | IRGQ         | MIR4784     | NIT2        | PPP1R7      | RTF1        | TACO1       | VAMP7       |
| AP1S1       | C12orf10    | CXorf40B    | FDPS        | ITFG1        | MIR636      | NME1        | PPP2CA      | RTN4        | TCEA2       | VLDLR       |
| AP2M1       | C1orf216    | CYC1        | FECH        | ITPA         | MIR6890     | NOP16       | PPP2R1A     | RWDD2B      | TCEAL2      | VPS41       |
| AP2S1       | C21orf33    | CYP2E1      | FHIT        | ITPR1        | MLH1        | NRXN3       | PPP2R2B     | SAMM50      | TCEAL4      | VPS4B       |
| AP3M2       | C3orf62     | DCTN3       | FIBP        | KATNB1       | MMP24-AS1   | NSDHL       | PPP2R5D     | SARS        | TFCP2       | VPS51       |
| AP3S2       | C5orf30     | DCTN6       | FIG4        | KHDRBS1      | MOCS2       | NUDCD3      | PPP3CB      | SCAMP5      | TFPT        | WBP11       |
| APEH        | C6orf106    | DCTPP1      | FKBP1B      | KIAA0391     | MOSPD1      | NUDT2       | PRC1        | SCFD1       | TIMM10      | WDR44       |
| APEX1       | CCBL2       | DDA1        | FLRT1       | KIAA0513     | MPI         | NUP155      | PRKCZ       | SDHA        | TIMM10B     | WDR61       |
| APLP2       | CCDC51      | DDHD2       | FRMPD4      | KIAA1279     | MRPL15      | ODC1        | PRMT8       | SDHAF1      | TIMM8B      | WDR7        |
| APMAP       | CCK         | DDX1        | FTO         | KIF2A        | MRPL35      | ODF2        | PRPF4       | SDHAP1      | TMED3       | WDR77       |
| AREL1       | CCNH        | DDX24       | FUCA1       | KIF3C        | MRPL4       | OPTN        | PRR13       | SEC31A      | TMEM177     | WDYHV1      |
| ARF1        | CCSER2      | DDX42       | GABARAPL1   | KIFAP3       | MRPS18A     | ORC5        | PSD3        | SEH1L       | TMEM208     | YWHAZ       |
| ARF3        | CDC123      | DGUOK       | GABBR1      | KLHDC4       | MRPS22      | OSBP        | PSMA1       | SERINC3     | TMEM246     | ZDHHC4      |
| ARFGAP2     | CDC42       | DHRS11      | GALNT11     | L1CAM        | MRPS33      | OXCT1       | PSMA5       | SEZ6L2      | TMEM41B     | ZDHHC6      |
| ARHGEF9     | CDK20       | DHRS7B      | GARS        | LANCL1       | MRPS35      | OXLD1       | PSMB3       | SF3A3       | TMEM97      | ZMYM4       |
| ARL2BP      | CDK5        | DIABLO      | GBAS        | LANCL2       | MRPS7       | PAAF1       | PSMB4       | SF3B5       | TMUB2       | ZNF365      |
| ARMC2-AS1   | CDK7        | DKK3        | GGNBP2      | LBH          | MRTO4       | PAFAH1B1    | PSMB5       | SH3BP5      | TOMM20      | ZNF593      |
| ARPC1A      | CDS2        | DLD         | GLO1        | LCMT1        | MTCH1       | PARL        | PSMB6       | SIPA1L1     | TOR1AIP2    | ZNF629      |
| ARPC5L      | CERK        | DNAJC8      | GLOD4       | LDHA         | MTCH2       | PARN        | PSMB7       | SLC23A2     | TOX4        |             |
| ASS1        | CFL1        | DRG1        | GLT8D1      | LDHB         | MTCL1       | PCCB        | PSMC1       | SLC25A11    | TP53BP1     |             |
| ASTN2       | CHCHD2      | DSTN        | GMPR2       | LDLRAD4      | MTHFD1      | PCYOX1L     | PSMC2       | SLC25A3     | TPGS2       |             |
| ATG14       | CHP1        | DUSP26      | GNB1        | LDOC1        | MTPAP       | PDCL3       | PSMC5       | SLC25A46    | TRAP1       |             |
| ATP1A1      | CHST1       | DYNC1H1     | GNB5        | LETMD1       | MTX2        | PDHB        | PSMD1       | SLC25A5     | TRAPPC2L    |             |
| ATP1A3      | CIAPIN1     | DYNC1I1     | GOT1        | LINC00094    | MX1         | PDK2        | PSMD11      | SLC25A6     | TSG101      |             |
| ATP2B2      | CIRBP       | DZIP3       | GOT2        | LMO3         | MYH10       | PDZD8       | PSMD14      | SLC30A9     | TSPAN3      |             |
| ATP5A1      | CITED1      | EAPP        | GPD1L       | LOC101927673 | MYL12B      | PEX14       | PSMD2       | SLC35B1     | TSPAN7      |             |
| ATP5B       | CLCN6       | EIF2AK1     | GPI         | LOC645166    | MYL5        | PEX19       | PSMD7       | SLC6A1      | TSR2        |             |


## Mouse data

Mouse data was sourced from GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE4758, referenced as CON, however during the analysis I realised that alpha synuclein had a positive log fold change, even though every other dataset reported a negative log fold change for sporadic parkinson's disease. Looking back at the dataset I realised it was an overexpression model rather than a mutant. Woops!

## Alpha synuclein expression

What's interesting is that canonically, alpha synuclein has been suggested to be overexpressed in sporadic parkinson's patients. 

#Filtering out other signals

As was done with the TDP-43 pipeline, I had to find other signals that could be contaminating the sporadic signal. I decided to use the ALS datasets from the TDP-43 analysis to represent "general neurodegeneration", (C9orf72, sALS, PET, RAV, SOD1, FUS) but actually there was very little overlap:

### Common Up (ALS)
ALOX5
ALPK1
STOM
MAFF
COL7A1

### Common Down (ALS)
PMS2P1
ATP6V1G2
INSIG1
RRAGA
TPGS2
MTCH2
PPP2CA
GLO1
PDHB


In a similar vein I decided to source some sporadic Parkinson's disease blood - to make sure it was leaving a more tissue specific signal. I found two datasets on GEO - GSE99039 (AMA) and GSE72267 (RON). These were bigger datasets than I usually use - AMA was total 353 samples, RON 59. What I realised was that my filters for the presence calls were still 2. This meant that 24,000 genes/probes passed through filtering which was making huge overlaps. I realised a filter of 2 wasn't really proportional to the number of samples in AMA, so I increased this to 20 (as the dataset is approximately 10 times larger than the other datasets I have been using).

These two datasets produced a common gene expression signature of approximately 5000 genes, which is quite a lot of concordance. When you compare this to the brain signal, there is an overlap of 88 upregulated and 76 downregulated genes. This is quite a small overlap considering the size of the blood signature. 

At this point, the remaining gene list for the sporadic signature is 558 genes (look for *Filt_ALS_blood* files)

## Enrichr

Down http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=3rwmj

Up http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=3rwmr

### CON DATA

There was a dataset I labelled "CON" with the GSEnumber GSE4758 that I included for a little while as it was a mouse model of alpha synuclein which I later removed as the research question changed.

### Adding LRRK2 Data

Oliver Bandmann mentioned that there is a phenomenon that LRRK2 patients exhibit their disease more closely to sporadic patients than other familial patients. I thought this was interesting and could be a good application of my methdology.

I sourced two datasets with LRRK2 patient data - GSE34516 and GSE23290. These were exon array so had to first be preprocessed by Wenbin in the Affymetrix Console using RMA-sketch. The output was normalised expression data. Mas5Calls didn't seem to work so I skipped this step. 

```{r, eval = F}
#### BOT GSE34516 #########
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Data/")
Data <- read.csv("GSE34516.RMA-GENE-EXTENDED-EC-hg19-na36_LRRK2only.csv", row.names = 1)

genename <- as.data.frame(Data$GeneSymbol)
rownames(genename) <- rownames(Data)
colnames(genename) <- "Gene.Symbol"
#Data is already normalised using RMA-sketch
analysis.name<-"BOT" #Label analysis
expressionMatrix<-Data[,1:6] #takes expression from normalised expression set


Treat<-factor(rep(c("Control", "Patient"),c(4,2)), levels=c("Control", "Patient"))
design<-model.matrix(~Treat)
rownames(design)<-colnames(expressionMatrix)
design

#Conduct statistical analysis of expression
library(limma)
fit<-lmFit(expressionMatrix, design) #linear model fit
fit<-eBayes(fit) 
result<-topTable(fit, coef="TreatPatient", adjust="BH", number=nrow(expressionMatrix)) #"BH" adjust for multiple hypothesis testing
#toptable normally takes top number but this takes all
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression")
write.csv(result, file = paste(analysis.name, "_result.csv", sep = ""))


result$"Fold Change"<-2^result$logFC 
result$"Fold Change"[result$"Fold Change"<1]<-(-1)/result$"Fold Change"[result$"Fold Change"<1] #converts log fold change into a linear value above or below 0
expressionLinear<-as.data.frame(2^expressionMatrix)
expressionLinear$ProbeSetID<-rownames(expressionLinear)
result<-merge(result, expressionLinear, by = 0) #merge values into one array
result<-merge(result, genename, by.x = "Row.names", by.y = 0)
result<-subset(result, subset=(Gene.Symbol !="---")) #if no gene symbol, discount


setwd(dir = "/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
write.csv(result, file=paste(analysis.name, "_finalresult.csv", sep=""), row.names=FALSE, quote = FALSE)

genesort <- result[order(result$P.Value),]
uniqueresult <- genesort[!duplicated(genesort[,16]),]
write.csv(uniqueresult, file=paste(analysis.name, "rankeduniqueresult.csv", sep=""), row.names=FALSE, quote = FALSE)

#### BOT2 GSE23290 #########
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Data/")
Data <- read.csv("GSE23290.RMA-GENE-EXTENDED-EC-hg19-na36_LRRK2only.csv", row.names = 1)

genename <- as.data.frame(Data$GeneSymbol)
rownames(genename) <- rownames(Data)
colnames(genename) <- "Gene.Symbol"
#Data is already normalised using RMA-sketch
analysis.name<-"BOT2" #Label analysis
expressionMatrix<-Data[,1:8] #takes expression from normalised expression set


Treat<-factor(rep(c("Control", "Patient"),c(5,3)), levels=c("Control", "Patient"))
design<-model.matrix(~Treat)
rownames(design)<-colnames(expressionMatrix)
design

#Conduct statistical analysis of expression
library(limma)
fit<-lmFit(expressionMatrix, design) #linear model fit
fit<-eBayes(fit) 
result<-topTable(fit, coef="TreatPatient", adjust="BH", number=nrow(expressionMatrix)) #"BH" adjust for multiple hypothesis testing
#toptable normally takes top number but this takes all
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression")
write.csv(result, file = paste(analysis.name, "_result.csv", sep = ""))


result$"Fold Change"<-2^result$logFC 
result$"Fold Change"[result$"Fold Change"<1]<-(-1)/result$"Fold Change"[result$"Fold Change"<1] #converts log fold change into a linear value above or below 0
expressionLinear<-as.data.frame(2^expressionMatrix)
expressionLinear$ProbeSetID<-rownames(expressionLinear)
result<-merge(result, expressionLinear, by = 0) #merge values into one array
result<-merge(result, genename, by.x = "Row.names", by.y = 0)
result<-subset(result, subset=(Gene.Symbol !="---")) #if no gene symbol, discount


setwd(dir = "/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
write.csv(result, file=paste(analysis.name, "_finalresult.csv", sep=""), row.names=FALSE, quote = FALSE)

genesort <- result[order(result$P.Value),]
uniqueresult <- genesort[!duplicated(genesort[,18]),]
write.csv(uniqueresult, file=paste(analysis.name, "rankeduniqueresult.csv", sep=""), row.names=FALSE, quote = FALSE)

```

I had to skip out a number of steps because it didn't require the annotation file. There were about 28000 rows left over which was quite a lot, though approximately 7000 were LOC transcripts so took us down to the number of genes you would expect. 


After adding these two datasets to the FoldChange_PD.R script, and filtering out for ALS and sporadic PD blood, there were 85 upregulated and 109 downregulated genes. This still included MSX1, SNCA, UCHL1, and ALDH1A1 from PD Malacards but not PINK1 anymore.

EnrichR
UP: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=3t9ff
DOWN: http://amp.pharm.mssm.edu/Enrichr/enrich?dataset=3t9fh


# Change of direction

The other datasets I had were fibroblasts from PARK2 and LRRK2 patients as provided by Robin Highly/Oliver Bandmann/Heather Mortiboys. These were exon array datasets which were analysed by Wenbin using the Affymetrix console. When these were added to the filters, the number of common genes was 36, only one of which MSX1 was retained. 

To give a little more signal, but to also see if the signal could be recovered, I discluded the blood data (as fibroblast is representative enough of a non-neuronal tissue), leaving 54 genes

DOCK2
RAD52
TNFAIP8
P4HA1
MS4A6A
CD300A
KLHL36
TRIM5
CD163
CXCR4
RIPK2
PAWR
ZNF516
JMJD6
CASP7
MSX1
IGF1R
LRRC1
SAFB2
C1QB
IL17RB
LAMB2
CCDC40
CDA
FMO5
DNASE2
NFE2
ZNF217
HMOX1
PARP16
HP1BP3
PGPEP1
CD14
NPIPB15
IMP4
LETMD1
AP1S1
FHIT
MECR
ATP2B2
DCTPP1
WDYHV1
PSMC2
NRXN3
MRTO4
NDUFS3
SLC6A1
NME1
MRPL15
COPS3
TUBG1
PRC1
BFSP1
IARS


Approximately 10 of these genes have a relatively strong relationship with Parkinson's or LRRK2 specifically. 32 were directly related according to VarElect, 22 indirectly related (VarElect_Results_cmgreen1-sheffield-ac-uk-20180522-014621619).


The next step was to find the PPI partners of these 54 genes.

# PPI 
```{r, eval=F}
#### ALS PARK2 LRRK2 ####

library(biomaRt)

setwd("/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/PPI_Network/")

PPI <- read.table("iref14_Human_UP_noDup_table_nodash.txt", header = T)
braingenes <- read.csv("Zhang_BrainCelltype_Markers_braingenes.csv", header = T)

DEG_list <- readLines("~/Documents/PhD/Parkinsons/Parkinsons_Code/Results/ALSPARK2LRRK2/removeallALLgenes.txt")

mart <- useMart("ENSEMBL_MART_ENSEMBL",dataset="hsapiens_gene_ensembl", host="www.ensembl.org")
attributes <- listAttributes(mart)
mart_back <- getBM(attributes =c("hgnc_symbol", "uniprotswissprot"), filters="hgnc_symbol", values=DEG_list,  mart=mart)
genelist_Uniprot <- subset(mart_back, !(mart_back$uniprotswissprot == ""))
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/ALSPARK2LRRK2/")
write.csv(genelist_Uniprot, "martback.csv", row.names = F)

# IDENTIFY MISSING GENES AND FIND UNIPROT CODES FOR THEM. NB SOME GENES MAY NOT BE PROTEIN CODING #

mart_table <- read.csv("martback.csv", header = T) #A table with the uniprot codes for the DEGs
uniprot_gene <- mart_table$uniprotswissprot

DEG_PPI <- subset(PPI, PPI$V1 %in% uniprot_gene | PPI$V2 %in% uniprot_gene)
rownames(DEG_PPI) <- 1:nrow(DEG_PPI)

write.csv(DEG_PPI, "DEG_PPI_ALP.csv", row.names = F)

DEG_PPI <- read.csv("DEG_PPI_ALP.csv")

DEG_PPI <- subset(DEG_PPI, DEG_PPI$Gene1 !="-")
DEG_PPI <- subset(DEG_PPI, DEG_PPI$Gene2 !="-")

write.csv(DEG_PPI, "FinalPDPPI.csv", row.names = F, quote = F)
```

From this we have a network of 1367 interactions between 1034 proteins (file = PD_DEGPPI.R, DEG_PPI_ALP.csv, ALP_PPIgenes.txt. Gene names used).  

Next I had to find the corresonding rows in each of the RNA expression data files ready for the correlation anaylysis. The LRRK2 samples were not included in the coexpression analysis because with only 2 and 3 patients respectively, it would results in rho values of 1/-1 or 1, 0.5, 0, -0.5 or 1, which are difficult to threshold. I also removed the MOR.FC dataset as the distribution of Rho values was abnormally distributed

[]

```{r, eval=F}
#### ALP####
#Read in network nodes
DEG_PPI <- readLines("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/ALSPARK2LRRK2/ALP_PPIGenes.txt")

setwd("/users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
#Extract PPI network genes from each dataset#####
LEW <- read.csv("LEWfilteredresult.csv")
rownames(LEW) <- LEW$Gene.Symbol
LEW <- LEW[,28:41]
LEW <- subset(LEW, rownames(LEW) %in% DEG_PPI)

MID3 <- read.csv("MID3filteredresult.csv")
rownames(MID3) <- MID3$Gene.Symbol
MID3 <- MID3[,36:49]
MID3 <- subset(MID3, rownames(MID3) %in% DEG_PPI)

MID4 <- read.csv("MID4filteredresult.csv")
rownames(MID4) <- MID4$Gene.Symbol
MID4 <- MID4[,41:55]
MID4 <- subset(MID4, rownames(MID4) %in% DEG_PPI)

MOR.FC <- read.csv("MOR.FCfilteredresult.csv")
rownames(MOR.FC) <- MOR.FC$Gene.Symbol
MOR.FC <- MOR.FC[,24:28]
MOR.FC <- subset(MOR.FC, rownames(MOR.FC) %in% DEG_PPI)

DIJ <- read.csv("DIJfilteredresult.csv")
rownames(DIJ) <- DIJ$Gene.Symbol
DIJ <- DIJ[,29:43]
DIJ <- subset(DIJ, rownames(DIJ) %in% DEG_PPI)

FFR <- read.csv("FFRfilteredresult.csv")
rownames(FFR) <- FFR$Gene.Symbol
FFR <- FFR[,29:44]
FFR <- subset(FFR, rownames(FFR) %in% DEG_PPI)

MID1 <- read.csv("MID1filteredresult.csv")
rownames(MID1) <- MID1$Gene.Symbol
MID1 <- MID1[,28:37]
MID1 <- subset(MID1, rownames(MID1) %in% DEG_PPI)

MID2 <- read.csv("MID2filteredresult.csv")
rownames(MID2) <- MID2$Gene.Symbol
MID2 <- MID2[,39:49]
MID2 <- subset(MID2, rownames(MID2) %in% DEG_PPI)

MOR.SN <- read.csv("MOR.SNfilteredresult.csv")
rownames(MOR.SN) <- MOR.SN$Gene.Symbol
MOR.SN <- MOR.SN[,36:59]
MOR.SN <- subset(MOR.SN, rownames(MOR.SN) %in% DEG_PPI)

DUM <- read.csv("DUM_UniqueGene_DESeq2.csv")
rownames(DUM) <- DUM$hgnc_symbol
DUM <- DUM[,53:81]
DUM <- subset(DUM, rownames(DUM) %in% DEG_PPI)




#Find the gene names that all datasets have in common
DEG_com <- Reduce(intersect, list(rownames(DIJ), rownames(FFR), 
                                  rownames(LEW),rownames(MID1),rownames(MID2),
                                  rownames(MID3),rownames(MID4),rownames(MOR.FC),
                                  rownames(MOR.SN), rownames(DUM)))
```

This script resulted in 854 common names. Correlation analysis was conducted on sharc (/shared/hidelab2/user/mdp15cmg/Parkinsons/PD_Cor)

When the correlation relationships were downloaded and analysed, only 59 edges reached a consistent correlation value of over 0.1 or below -0.1 regardless of sign. This increases to 208 when the RNA-seq dataset is removed. 


# Yet Another New Approach

Talking to Kat we hypothesised that the Parkin Fibroblasts were perhaps driving away the signal too strongly. Also the number of samples was so small that it's not necessarily giving a representative signal. 

To compensate, I decided to go back to the blood samples and use some other samples from the AMA GSE99039 dataset which have familial mutations. There were 5 ATP13A2 cases, 15 Parkin cases, and 12 PINK1 cases. Two of the Parkin cases in the metadata didn't have corresponding files, so this became 13 Parkin cases. 

## Limma
```{r, eval=f}
##### AMA ATP13A2 #########
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Data/GSE99039_keepfiles/")

#run program to choose .CEL files from directory
celfiles <- fileBrowser(textToShow = "Choose CEL files", testFun = hasSuffix("[cC][eE][lL]"))
Data<-ReadAffy(filenames=celfiles) #read in files
rmaEset<-rma(Data) #normalise using RMA
analysis.name<-"AMA_ATP13A2" #Label analysis
dataMatrixAll<-exprs(rmaEset) #takes expression from normalised expression set


#mas5call generates presence/absence calls for each probeset
mas5call<-mas5calls(Data)
callMatrixAll<-exprs(mas5call)
colnames(callMatrixAll)<-sub(".CEL", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
colnames(callMatrixAll)<-sub(".cel", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
callMatrixAll<-as.data.frame(callMatrixAll)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPf<-function(x){
  sum(x=="P")
}

#count how many samples have presence calls
countPl<-apply(callMatrixAll, 1, countPf)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPdf<-data.frame(ProbeSetID=names(countPl), countP=countPl) 

#read annotation
# USING ANNOTATION FILE (if .csv, convert to .txt using excel)
annotation.file<-"/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/HG-U133_Plus_2.na35.annot.csv/HG-U133_Plus_2.na35_SHORT.annot.txt"
annotation<-read.table(annotation.file, header = TRUE, row.names=NULL, sep="\t", skip=0, stringsAsFactors=F, quote = "", comment.char="!", fill = TRUE, as.is = TRUE)
dim(annotation)
nrow(annotation)
annotation<-subset( annotation, subset=(Gene.Symbol !="---")) #if no gene symbol, discount

# Remove rows in which genes are noted to have negative strand matching probes
idxNegativeStrand<-grep("Negative Strand Matching Probes", annotation$Annotation.Notes)
if(length(idxNegativeStrand)>0)
{
  annotation<-annotation[-idxNegativeStrand,]
}


expressionMatrix<-exprs(rmaEset)
colnames(expressionMatrix)

#this is for matched samples
Treat<-factor(rep(c("Control", "Patient"),c(183,5)), levels=c("Control", "Patient"))
design<-model.matrix(~Treat)
rownames(design)<-colnames(expressionMatrix)
design

#Conduct statistical analysis of expression
library(limma)
fit<-lmFit(expressionMatrix, design) #linear model fit
fit<-eBayes(fit) 
result<-topTable(fit, coef="TreatPatient", adjust="BH", number=nrow(expressionMatrix)) #"BH" adjust for multiple hypothesis testing
#toptable normally takes top number but this takes all
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression")
write.csv(result, file = paste(analysis.name, "_result.csv", sep = ""))

result$"ProbeSetID"<-rownames(result) #make probeset IDs the row names
head(result$"ProbeSetID") 
result$"Fold Change"<-2^result$logFC 
result$"Fold Change"[result$"Fold Change"<1]<-(-1)/result$"Fold Change"[result$"Fold Change"<1] #converts log fold change into a linear value above or below 0
expressionLinear<-as.data.frame(2^expressionMatrix)
expressionLinear$ProbeSetID<-rownames(expressionLinear)
result<-merge(result, expressionLinear, by.x="ProbeSetID", by.y="ProbeSetID") #merge values into one array
result<-merge(annotation, result, by.x="Probe.Set.ID", by.y="ProbeSetID")
result<-merge(result, countPdf, by.x="Probe.Set.ID", by.y="ProbeSetID")
result$Gene.Symbol <- sapply(strsplit(result$Gene.Symbol,"///"), `[`, 1)
result$Ensembl <- sapply(strsplit(result$Ensembl,"///"), `[`, 1)

setwd(dir = "/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
write.csv(result, file=paste(analysis.name, "_finalresult.csv", sep=""), row.names=FALSE, quote = FALSE)

genesort <- result[order(result$P.Value),]
uniqueresult <- genesort[!duplicated(genesort[,5]),]
write.csv(uniqueresult, file=paste(analysis.name, "rankeduniqueresult.csv", sep=""), row.names=FALSE, quote = FALSE)

uniqueresult<-subset(uniqueresult, Gene.Symbol!="") #removes any probes for which there are no gene symbols
uniqueresult<-subset(uniqueresult, subset=(countP>20)) #only takes results that have at least 2 samples with a presence call for a probe
write.csv(uniqueresult, file=paste(analysis.name, "filteredresult.csv", sep=""), row.names=FALSE, quote = FALSE)

##### AMA PINK1 ##########
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Data/GSE99039_keepfiles/")

#run program to choose .CEL files from directory
celfiles <- fileBrowser(textToShow = "Choose CEL files", testFun = hasSuffix("[cC][eE][lL]"))
Data<-ReadAffy(filenames=celfiles) #read in files
rmaEset<-rma(Data) #normalise using RMA
analysis.name<-"AMA_PINK1" #Label analysis
dataMatrixAll<-exprs(rmaEset) #takes expression from normalised expression set


#mas5call generates presence/absence calls for each probeset
mas5call<-mas5calls(Data)
callMatrixAll<-exprs(mas5call)
colnames(callMatrixAll)<-sub(".CEL", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
colnames(callMatrixAll)<-sub(".cel", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
callMatrixAll<-as.data.frame(callMatrixAll)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPf<-function(x){
  sum(x=="P")
}

#count how many samples have presence calls
countPl<-apply(callMatrixAll, 1, countPf)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPdf<-data.frame(ProbeSetID=names(countPl), countP=countPl) 

#read annotation
# USING ANNOTATION FILE (if .csv, convert to .txt using excel)
annotation.file<-"/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/HG-U133_Plus_2.na35.annot.csv/HG-U133_Plus_2.na35_SHORT.annot.txt"
annotation<-read.table(annotation.file, header = TRUE, row.names=NULL, sep="\t", skip=0, stringsAsFactors=F, quote = "", comment.char="!", fill = TRUE, as.is = TRUE)
dim(annotation)
nrow(annotation)
annotation<-subset( annotation, subset=(Gene.Symbol !="---")) #if no gene symbol, discount

# Remove rows in which genes are noted to have negative strand matching probes
idxNegativeStrand<-grep("Negative Strand Matching Probes", annotation$Annotation.Notes)
if(length(idxNegativeStrand)>0)
{
  annotation<-annotation[-idxNegativeStrand,]
}


expressionMatrix<-exprs(rmaEset)
colnames(expressionMatrix)

#this is for matched samples
Treat<-factor(rep(c("Control", "Patient"),c(183,12)), levels=c("Control", "Patient"))
design<-model.matrix(~Treat)
rownames(design)<-colnames(expressionMatrix)
design

#Conduct statistical analysis of expression
library(limma)
fit<-lmFit(expressionMatrix, design) #linear model fit
fit<-eBayes(fit) 
result<-topTable(fit, coef="TreatPatient", adjust="BH", number=nrow(expressionMatrix)) #"BH" adjust for multiple hypothesis testing
#toptable normally takes top number but this takes all
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression")
write.csv(result, file = paste(analysis.name, "_result.csv", sep = ""))

result$"ProbeSetID"<-rownames(result) #make probeset IDs the row names
head(result$"ProbeSetID") 
result$"Fold Change"<-2^result$logFC 
result$"Fold Change"[result$"Fold Change"<1]<-(-1)/result$"Fold Change"[result$"Fold Change"<1] #converts log fold change into a linear value above or below 0
expressionLinear<-as.data.frame(2^expressionMatrix)
expressionLinear$ProbeSetID<-rownames(expressionLinear)
result<-merge(result, expressionLinear, by.x="ProbeSetID", by.y="ProbeSetID") #merge values into one array
result<-merge(annotation, result, by.x="Probe.Set.ID", by.y="ProbeSetID")
result<-merge(result, countPdf, by.x="Probe.Set.ID", by.y="ProbeSetID")
result$Gene.Symbol <- sapply(strsplit(result$Gene.Symbol,"///"), `[`, 1)
result$Ensembl <- sapply(strsplit(result$Ensembl,"///"), `[`, 1)

setwd(dir = "/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
write.csv(result, file=paste(analysis.name, "_finalresult.csv", sep=""), row.names=FALSE, quote = FALSE)

genesort <- result[order(result$P.Value),]
uniqueresult <- genesort[!duplicated(genesort[,5]),]
write.csv(uniqueresult, file=paste(analysis.name, "rankeduniqueresult.csv", sep=""), row.names=FALSE, quote = FALSE)

uniqueresult<-subset(uniqueresult, Gene.Symbol!="") #removes any probes for which there are no gene symbols
uniqueresult<-subset(uniqueresult, subset=(countP>20)) #only takes results that have at least 2 samples with a presence call for a probe
write.csv(uniqueresult, file=paste(analysis.name, "filteredresult.csv", sep=""), row.names=FALSE, quote = FALSE)

##### AMA PARKIN ##########
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Data/GSE99039_keepfiles/")

#run program to choose .CEL files from directory
celfiles <- fileBrowser(textToShow = "Choose CEL files", testFun = hasSuffix("[cC][eE][lL]"))
Data<-ReadAffy(filenames=celfiles) #read in files
rmaEset<-rma(Data) #normalise using RMA
analysis.name<-"AMA_PARKIN" #Label analysis
dataMatrixAll<-exprs(rmaEset) #takes expression from normalised expression set


#mas5call generates presence/absence calls for each probeset
mas5call<-mas5calls(Data)
callMatrixAll<-exprs(mas5call)
colnames(callMatrixAll)<-sub(".CEL", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
colnames(callMatrixAll)<-sub(".cel", ".mas5-Detection", colnames(callMatrixAll),fixed=TRUE)
callMatrixAll<-as.data.frame(callMatrixAll)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPf<-function(x){
  sum(x=="P")
}

#count how many samples have presence calls
countPl<-apply(callMatrixAll, 1, countPf)
callMatrixAll$ProbeSetID<-rownames(callMatrixAll)
countPdf<-data.frame(ProbeSetID=names(countPl), countP=countPl) 

#read annotation
# USING ANNOTATION FILE (if .csv, convert to .txt using excel)
annotation.file<-"/Users/clairegreen/Documents/PhD/TDP-43/TDP-43_Data/HG-U133_Plus_2.na35.annot.csv/HG-U133_Plus_2.na35_SHORT.annot.txt"
annotation<-read.table(annotation.file, header = TRUE, row.names=NULL, sep="\t", skip=0, stringsAsFactors=F, quote = "", comment.char="!", fill = TRUE, as.is = TRUE)
dim(annotation)
nrow(annotation)
annotation<-subset( annotation, subset=(Gene.Symbol !="---")) #if no gene symbol, discount

# Remove rows in which genes are noted to have negative strand matching probes
idxNegativeStrand<-grep("Negative Strand Matching Probes", annotation$Annotation.Notes)
if(length(idxNegativeStrand)>0)
{
  annotation<-annotation[-idxNegativeStrand,]
}


expressionMatrix<-exprs(rmaEset)
colnames(expressionMatrix)

#this is for matched samples
Treat<-factor(rep(c("Control", "Patient"),c(183,13)), levels=c("Control", "Patient"))
design<-model.matrix(~Treat)
rownames(design)<-colnames(expressionMatrix)
design

#Conduct statistical analysis of expression
library(limma)
fit<-lmFit(expressionMatrix, design) #linear model fit
fit<-eBayes(fit) 
result<-topTable(fit, coef="TreatPatient", adjust="BH", number=nrow(expressionMatrix)) #"BH" adjust for multiple hypothesis testing
#toptable normally takes top number but this takes all
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression")
write.csv(result, file = paste(analysis.name, "_result.csv", sep = ""))

result$"ProbeSetID"<-rownames(result) #make probeset IDs the row names
head(result$"ProbeSetID") 
result$"Fold Change"<-2^result$logFC 
result$"Fold Change"[result$"Fold Change"<1]<-(-1)/result$"Fold Change"[result$"Fold Change"<1] #converts log fold change into a linear value above or below 0
expressionLinear<-as.data.frame(2^expressionMatrix)
expressionLinear$ProbeSetID<-rownames(expressionLinear)
result<-merge(result, expressionLinear, by.x="ProbeSetID", by.y="ProbeSetID") #merge values into one array
result<-merge(annotation, result, by.x="Probe.Set.ID", by.y="ProbeSetID")
result<-merge(result, countPdf, by.x="Probe.Set.ID", by.y="ProbeSetID")
result$Gene.Symbol <- sapply(strsplit(result$Gene.Symbol,"///"), `[`, 1)
result$Ensembl <- sapply(strsplit(result$Ensembl,"///"), `[`, 1)

setwd(dir = "/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")
write.csv(result, file=paste(analysis.name, "_finalresult.csv", sep=""), row.names=FALSE, quote = FALSE)

genesort <- result[order(result$P.Value),]
uniqueresult <- genesort[!duplicated(genesort[,5]),]
write.csv(uniqueresult, file=paste(analysis.name, "rankeduniqueresult.csv", sep=""), row.names=FALSE, quote = FALSE)

uniqueresult<-subset(uniqueresult, Gene.Symbol!="") #removes any probes for which there are no gene symbols
uniqueresult<-subset(uniqueresult, subset=(countP>20)) #only takes results that have at least 2 samples with a presence call for a probe
write.csv(uniqueresult, file=paste(analysis.name, "filteredresult.csv", sep=""), row.names=FALSE, quote = FALSE)
```

## Common Differentially Expressed Genes (Blood Edition)
```{r, eval=f}
#############################################################
################### FOLD CHANGE DEG #########################
#############################################################

setwd("/users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")

LEW <- read.csv("LEWfilteredresult.csv")
LEW <- LEW[order(LEW$P.Value),]
MID3 <- read.csv("MID3filteredresult.csv")
MID3 <- MID3[order(MID3$P.Value),]
MID4 <- read.csv("MID4filteredresult.csv")
MID4 <- MID4[order(MID4$P.Value),]
MOR.FC <- read.csv("MOR.FCfilteredresult.csv")
MOR.FC <- MOR.FC[order(MOR.FC$P.Value),]
DIJ <- read.csv("DIJfilteredresult.csv")
DIJ <- DIJ[order(DIJ$P.Value),]
FFR <- read.csv("FFRfilteredresult.csv")
FFR <- FFR[order(FFR$P.Value),]
MID1 <- read.csv("MID1filteredresult.csv")
MID1 <- MID1[order(MID1$P.Value),]
MID2 <- read.csv("MID2filteredresult.csv")
MID2 <- MID2[order(MID2$P.Value),]
MOR.SN <- read.csv("MOR.SNfilteredresult.csv")
MOR.SN <- MOR.SN[order(MOR.SN$P.Value),]
DUM <- read.csv("DUM_UniqueGene_DESeq2.csv")
DUM <- DUM[order(DUM$pvalue),]
BOT <- read.csv("BOTrankeduniqueresult.csv")
BOT <- BOT[order(BOT$P.Value),]
BOT2 <- read.csv("BOT2rankeduniqueresult.csv")
BOT2 <- BOT2[order(BOT2$P.Value),]




thresh <- 1

upLEW <- subset(LEW, LEW$Fold.Change >= thresh)
upLEWgene <- upLEW$Gene.Symbol

upMID3<- subset(MID3, MID3$Fold.Change >= thresh)
upMID3gene <- upMID3$Gene.Symbol

upMID4 <- subset(MID4, MID4$Fold.Change >= thresh)
upMID4gene <- upMID4$Gene.Symbol

upMOR.FC <- subset(MOR.FC, MOR.FC$Fold.Change >= thresh)
upMOR.FCgene <- upMOR.FC$Gene.Symbol

upDUM <- subset(DUM, DUM$log2FoldChange >= 0)
upDUMgene <- upDUM$hgnc_symbol

upDIJ <- subset(DIJ, DIJ$Fold.Change >= thresh)
upDIJgene <- upDIJ$Gene.Symbol

upFFR<- subset(FFR, FFR$Fold.Change >= thresh)
upFFRgene <- upFFR$Gene.Symbol

upMID1<- subset(MID1, MID1$Fold.Change >= thresh)
upMID1gene <- upMID1$Gene.Symbol

upMID2 <- subset(MID2, MID2$Fold.Change >= thresh)
upMID2gene <- upMID2$Gene.Symbol

upMOR.SN <- subset(MOR.SN, MOR.SN$Fold.Change >= thresh)
upMOR.SNgene <- upMOR.SN$Gene.Symbol

upBOT <- subset(BOT, BOT$Fold.Change >= thresh)
upBOTgene <- upBOT$Gene.Symbol

upBOT2 <- subset(BOT2, BOT2$Fold.Change >= thresh)
upBOT2gene <- upBOT2$Gene.Symbol


INTUP <- Reduce(intersect, list(upLEWgene, upMID3gene, upMID4gene, upMOR.FCgene,
                                upDIJgene, upFFRgene, upMID1gene, upMID2gene, upMOR.SNgene, upDUMgene, upBOTgene, upBOT2gene))



#### DOWN ####
thresh <- -1

downLEW <- subset(LEW, LEW$Fold.Change <= thresh)
downLEWgene <- downLEW$Gene.Symbol

downMID3 <- subset(MID3, MID3$Fold.Change <= thresh)
downMID3gene <- downMID3$Gene.Symbol

downMID4 <- subset(MID4, MID4$Fold.Change <= thresh)
downMID4gene <- downMID4$Gene.Symbol

downMOR.FC <- subset(MOR.FC, MOR.FC$Fold.Change <= thresh)
downMOR.FCgene <- downMOR.FC$Gene.Symbol

downDUM <- subset(DUM, DUM$log2FoldChange <= 0)
downDUMgene <- downDUM$hgnc_symbol

downDIJ <- subset(DIJ, DIJ$Fold.Change <= thresh)
downDIJgene <- downDIJ$Gene.Symbol

downFFR<- subset(FFR, FFR$Fold.Change <= thresh)
downFFRgene <- downFFR$Gene.Symbol

downMID1<- subset(MID1, MID1$Fold.Change <= thresh)
downMID1gene <- downMID1$Gene.Symbol

downMID2 <- subset(MID2, MID2$Fold.Change <= thresh)
downMID2gene <- downMID2$Gene.Symbol

downMOR.SN <- subset(MOR.SN, MOR.SN$Fold.Change <= thresh)
downMOR.SNgene <- downMOR.SN$Gene.Symbol


downBOT <- subset(BOT, BOT$Fold.Change <= thresh)
downBOTgene <- downBOT$Gene.Symbol

downBOT2 <- subset(BOT2, BOT2$Fold.Change <= thresh)
downBOT2gene <- downBOT2$Gene.Symbol

INTDOWN <- Reduce(intersect, list(downLEWgene, downMID3gene, downMID4gene, downMOR.FCgene,
                                  downDIJgene, downFFRgene, downMID1gene, downMID2gene, 
                                  downMOR.SNgene, downDUMgene, downBOTgene, downBOT2gene))




########################### COMMON GENES ##############################
all <- c(INTUP, INTDOWN)
setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/FoldChange")
# write.table(INTUP,"Sp_LRRK2_upDEGs.txt", row.names = F, col.names = F, quote = F)
# write.table(INTDOWN,"Sp_LRRK2_downDEGs.txt", row.names = F, col.names = F, quote = F)
# write.table(all, "Sp_LRRK2_allDEGs.txt", row.names = F, col.names = F, quote = F)

PDgenes <- readLines("/Users/clairegreen/Documents/PhD/Parkinsons/ParkinsonsDiseaseMalacards.txt")

intersect(INTUP, PDgenes)
intersect(INTDOWN, PDgenes)



####### ALS signature ##########

setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/noMedian/")

C9 <- read.csv("C9_unique.csv")
C9 <- C9[order(C9$P.Value),]
sals <- read.csv("sals_unique.csv")
sals <- sals[order(sals$P.Value),]

setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/TDP-43_DEseq2/")

pet <- read.csv("PET_results_keepfiltering.csv")
rav <- read.csv("RAV_results_keepfiltering.csv")

setwd("/users/clairegreen/Documents/PhD/TDP-43/TDP-43_Code/Results/GeneExpression/non-TDP/")

FUS <- read.csv("FUSrankeduniqueresult.csv")
FUS <- FUS[order(FUS$P.Value),]
SOD1 <- read.csv("SOD1rankeduniqueresult.csv")
SOD1 <- SOD1[order(SOD1$P.Value),]


thresh <- 1

upC9 <- subset(C9, C9$Fold.Change >= thresh)
upC9gene <- upC9$Gene.Symbol

upSALS <- subset(sals, sals$Fold.Change >= thresh)
upSALSgene <- upSALS$Gene.Symbol

upPET <- subset(pet, pet$FoldChange >= thresh)
upPETgene <- upPET$hgnc_symbol

upRAV <- subset(rav, rav$FoldChange >= thresh)
upRAVgene <- upRAV$hgnc_symbol

upFUS <- subset(FUS, FUS$Fold.Change >= thresh)
upFUSgene <- upFUS$Gene.Symbol

upSOD1 <- subset(SOD1, SOD1$Fold.Change >= thresh)
upSOD1gene <- upSOD1$Gene.Symbol

INTUP_ALS <- Reduce(intersect, list(upC9gene, upSALSgene, upPETgene, upRAVgene, upFUSgene, upSOD1gene))


#### DOWN 
thresh <- -1

downC9 <- subset(C9, C9$Fold.Change <= thresh)
downC9gene <- downC9$Gene.Symbol

downSALS <- subset(sals, sals$Fold.Change <= thresh)
downSALSgene <- downSALS$Gene.Symbol

downPET <- subset(pet, pet$FoldChange <= thresh)
downPETgene <- downPET$hgnc_symbol

downRAV <- subset(rav, rav$FoldChange <= thresh)
downRAVgene <- downRAV$hgnc_symbol

downFUS <- subset(FUS, FUS$Fold.Change <= thresh)
downFUSgene <- downFUS$Gene.Symbol

downSOD1 <- subset(SOD1, SOD1$Fold.Change <= thresh)
downSOD1gene <- downSOD1$Gene.Symbol

INTDOWN_ALS <- Reduce(intersect, list(downC9gene, downSALSgene, downPETgene, downRAVgene, downFUSgene, downSOD1gene))



##### COMMON GENES ###
upremove <- Reduce(intersect, list (INTUP, INTUP_ALS))
downremove <- Reduce(intersect, list(INTDOWN, INTDOWN_ALS))



###### REMOVE COMMON GENES ###
resultsup <- subset(INTUP, !(INTUP %in% upremove))
resultsdown <- subset(INTDOWN, !(INTDOWN %in% downremove))
results <- c(resultsup, resultsdown)

####### sPD Blood signature ##########

setwd("/users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")

AMA <- read.csv("AMAfilteredresult.csv")
AMA <- AMA[order(AMA$P.Value),]
RON <- read.csv("RONfilteredresult.csv")
RON <- RON[order(RON$P.Value),]


thresh <- 1

upAMA <- subset(AMA, AMA$Fold.Change >= thresh)
upAMAgene <- upAMA$Gene.Symbol

upRON <- subset(RON, RON$Fold.Change >= thresh)
upRONgene <- upRON$Gene.Symbol


INTUP_blood <- Reduce(intersect, list(upAMAgene, upRONgene))


#### DOWN ###
thresh <- -1

downAMA <- subset(AMA, AMA$Fold.Change <= thresh)
downAMAgene <- downAMA$Gene.Symbol

downRON <- subset(RON, RON$Fold.Change <= thresh)
downRONgene <- downRON$Gene.Symbol


INTDOWN_blood <- Reduce(intersect, list(downAMAgene, downRONgene))



##### COMMON GENES ###
upremove2 <- Reduce(intersect, list(INTUP, INTUP_blood))
downremove2 <- Reduce(intersect, list(INTDOWN, INTDOWN_blood))



##### REMOVE COMMON GENES ###
resultsup <- subset(resultsup, !(resultsup %in% upremove2))
resultsdown <- subset(resultsdown, !(resultsdown %in% downremove2))
results <- c(resultsup, resultsdown)




setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/FoldChange/")
# write.table(resultsup, "Sp_LRRK2_FiltALSPDblood_UPgenes.txt", quote = F, row.names = F, col.names = F)
# write.table(resultsdown, "Sp_LRRK2_FiltALSPDblood_DOWNgenes.txt", quote = F, row.names = F, col.names = F)
# write.table(results, "Sp_LRRK2_FiltALSPDblood_ALLgenes.txt", quote = F, row.names = F, col.names = F)
# cat(resultsup, sep="\n")

intersect(resultsup, PDgenes)
intersect(resultsdown, PDgenes)


####### fPD Blood signature ##########

setwd("/users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/GeneExpression/")

AMA_A <- read.csv("AMA_ATP13A2filteredresult.csv")
AMA_A <- AMA_A[order(AMA_A$P.Value),]
AMA_PARK <- read.csv("AMA_PARKINfilteredresult.csv")
AMA_PARK <- AMA_PARK[order(AMA_PARK$P.Value),]
AMA_PINK <- read.csv("AMA_PINK1filteredresult.csv")
AMA_PINK <- AMA_PINK[order(AMA_PINK$P.Value),]


thresh <- 1

upAMA_A <- subset(AMA_A, AMA_A$Fold.Change >= thresh)
upAMA_Agene <- AMA_A$Gene.Symbol

upAMA_PARK<- subset(AMA_PARK, AMA_PARK$Fold.Change >= thresh)
upAMA_PARKgene <- upAMA_PARK$Gene.Symbol

upAMA_PINK<- subset(AMA_PINK, AMA_PINK$Fold.Change >= thresh)
upAMA_PINKgene <- upAMA_PINK$Gene.Symbol


INTUP_fam_blood <- Reduce(intersect, list(upAMA_Agene, upAMA_PARKgene, upAMA_PINKgene))


#### DOWN ###
thresh <- -1

downAMA_A <- subset(AMA_A, AMA_A$Fold.Change <= thresh)
downAMA_Agene <- AMA_A$Gene.Symbol

downAMA_PARK<- subset(AMA_PARK, AMA_PARK$Fold.Change <= thresh)
downAMA_PARKgene <- downAMA_PARK$Gene.Symbol

downAMA_PINK<- subset(AMA_PINK, AMA_PINK$Fold.Change <= thresh)
downAMA_PINKgene <- downAMA_PINK$Gene.Symbol


INTDOWN_fam_blood <- Reduce(intersect, list(downAMA_Agene, downAMA_PARKgene, downAMA_PINKgene))



##### COMMON GENES ###
upremove3 <- Reduce(intersect, list(INTUP, INTUP_fam_blood))
downremove3 <- Reduce(intersect, list(INTDOWN, INTDOWN_fam_blood))



##### REMOVE COMMON GENES ###
resultsup <- subset(resultsup, !(resultsup %in% upremove3))
resultsdown <- subset(resultsdown, !(resultsdown %in% downremove3))
results <- c(resultsup, resultsdown)




setwd("/Users/clairegreen/Documents/PhD/Parkinsons/Parkinsons_Code/Results/FamilialBlood/")
write.table(resultsup, "ALS_sfblood_UPgenes.txt", quote = F, row.names = F, col.names = F)
write.table(resultsdown, "ALS_sfblood_DOWNgenes.txt", quote = F, row.names = F, col.names = F)
write.table(results, "ALS_sfblood_ALLgenes.txt", quote = F, row.names = F, col.names = F)
cat(resultsup, sep="\n")

intersect(resultsup, PDgenes)
intersect(resultsdown, PDgenes)
```

